### 2020-05-30 writeup/intro.tex
use squishlist
minor
wip

\begin{abstract}
When does multi-task and transfer learning outperform single-task learning?
In this work, we address this question by studying the test performance of predicting a particular task given multiple tasks using a commonly used hard parameter sharing architecture, in terms of bais-variance tradeoff.
In the case of high-dimensional linear regression, we provide a sharp analysis of the bias-variance tradeoff for multi-task and transfer learning estimators.
A key technical tool that we develop is the trace of $(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}$ for two random matrices $X_1$ and $X_2$ in the setting of two tasks.
	Based on the theory, we provide more precise interpretations of many empirical phenomena in multi-task and transfer learning.
	For example, we quantify the benefit of multi-task learning for reducing the amount of labeled data, which is a key finding in Taskonomy by Zamir et al.'18.
	Finally, we show practical implications of our theory for detecting and improving negative effects in image and text classification tasks.
\end{abstract}

@@

%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning has become a powerful paradigm to solve complex prediction tasks in computer vision \cite{chexnet17,ZSSGM18}, natural language processing \cite{GLUE,superglue} and numerous other areas \cite{ZY17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
By combining multiple information sources, it is possible to share new information among different sources in the same model \cite{C97}.
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
Intuitively, the performance of multi-task learning depends on the relationship of the information sources.
When the information sources are heterogeneous, negative transfer -- where multi-task learning performs worse than single-task learning -- has often been observed \cite{AP16,BS17}.
While numerous studies have sought to alleviate negative transfer when it occurs \cite{ZY17}, a rigorous understanding to the contributing causes of negative transfer has remained elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.
In this work, we develop technical tools to better understand when and why negative transfer occurs for learning multiple linear regression tasks in high dimensions. % for learning from multiple linear regression tasks.
We use the tools to show theoretical and practical implications for detecting and improving negative transfer.
%We consider a setting where the target task has limited labeled data and show
%On the other hand, unless the structures across task data are well-understood, applying multi-task learning on several different datasets often result in suboptimal models (or negative transfer in more technical terms).

@@

%The technical challenge to develop a theory for multi-task learning is how to capture generalization performance that scales tightly with the amount of labeled data, in particular when the size of the training set is small.
%Prior generalization theory using uniform stability \cite{LTSM16}, Rademacher complexity \cite{BS03,BBCKP10} is unable to explain the phenomenon of negative transfer because there is no tight bound on test error that scales with the amount of labeled data.
%In particular in Figure \ref{fig_motivation}, we observe a shift from positive transfer to negative transfer as a parameter of task relatedness.
%The theory we develop will provide a precise explanation to such a phenomenon.

%To gain insight into the working of multi-task learning, we consider a simplified setting for learning multiple high-dimensional linear regression tasks.
%A typical process to do multi-task learning involves two steps:
%(i) Jointly learn a shared representation for all the tasks;
%(ii) Fine-tune the learnt model on a specific target task.
%We focus on a hard parameter sharing model proposed in \cite{R17,WZR20} and identify conditions on when multi-task and transfer learning works, and when it doesn't.
%The high-dimensional linear regression setting where the target task data size is limited captures the intuition that the target task only contains limited labeled data.

@@

We combine all the labeled data using a hard parameter sharing architecture that contains a shared body $B\in\real^{p\times r}$ for all tasks and a separate prediction head $\set{W_i \in \real^{r}}_{i=1}^t$ for each task \cite{R17,MTDNN19,WZR20}.
This corresponds to minimizing the following objective.
\begin{align}
	\label{eq_mtl}
	f(B; W_1, \dots, W_t) = \sum_{i=1}^t \norm{X_i B W_i - Y_i}^2.
\end{align}
For a target task $t$,
let $\hat{\beta}_t^{\MTL} = B W_t$ denote the optimal multi-task estimator by solving equation \eqref{eq_mtl}.
Let $\hat{\beta}_t^{\STL}$ denote the standard linear regression estimator using task $t$ alone.
We say there is negative transfer if the test error of $\hat{\beta}_t^{\MTL}$  is smaller than that of $\hat{\beta}_t^{\STL}$, or positive transfer otherwise.

@@

The tradeoff that we developed in Section \ref{sec_main} depends on the qualities of each task dataset such as its data size and covariance matrix.
In Section \ref{sec_insight}, we use the technical results to provide guidance on when positive transfer is likely to occur and identify causes of negative transfer.
\begin{itemize}
		\item First, we provide a more precise interpretation to a folklore understanding in multi-task learning, which posits that negative transfer is caused by having dissimilar tasks.
		For an example of two tasks, we show a sharp transition from positive to negative transfer determined by the ratio of $\norm{\beta_1 - \beta_2}^2$ and a certain function of data sizes (cf. Proposition \ref{prop_dist_transition}).
		We further show that positive transfer is more likely to occur when transferring from a less noisy source task to a more noisy target task.
		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
		\item Second, we find that depending on how large $\norm{\beta_1 - \beta_2}^2$ is, adding more labeled data from the source task does not always reduce the test error of $\hat{\beta}_t^{\MTL}$ (cf. Proposition \ref{prop_data_size}).
	We connect the observation to explain a key finding of Taskonomy \cite{ZSSGM18}.
	We define the \textit{data efficiency ratio} as the smallest $\alpha\in(0, 1)$ such that if we only use an $\alpha$ fraction of labeled data, then the test error of $\hat{\beta}_t^{\MTL}$ matches that of the $\hat{\beta}_t^{\STL}$ on the entire set.
	In Proposition \ref{prop_data_efficiency}, we show that the data efficiency ratio of an illustrative example is at most $\alert{\frac {1}{2\rho_t}}$.
	In Section \ref{sec_validate}, we validate that performing multi-task learning can reduce the need for labeled data on 6 sentiment analysis tasks. %, we observe that just by using \alert{40\%} of the labeled data, the overall accuracy of multi-task learning matches that of learning every task in isolation.
		\item Finally, we find that covariate shift, i.e. having different covariance matrices, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
		We show that as $n_1 / n_2$ becomes large, the best performing source task has have the same covariance matrix as the target task (cf. Proposition \ref{prop_covariate}).
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.
\end{itemize}

>>>>

\begin{abstract}
When does multi-task learning outperform single-task learning?
In this work, we address this question by studying the test performance of predicting a particular task given multiple tasks using a commonly used hard parameter sharing architecture, in terms of bais-variance tradeoff.
In the setting of high-dimensional linear regression, we provide a sharp analysis of the bias-variance tradeoff for multi-task learning estimators.
A key technical tool that we develop is the trace of $(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}$ for two random matrices $X_1$ and $X_2$ in the setting of two tasks.
	Based on the theory, we provide more precise interpretations of many empirical phenomena in multi-task learning.
	For example, we quantify the benefit of multi-task learning for reducing the amount of labeled data, which is a key finding in Taskonomy by Zamir et al.'18.
	Finally, we show practical implications of our theory for detecting and improving negative effects in image and text classification tasks.
\end{abstract}

@@

%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning represents a powerful paradigm to solve complex prediction tasks in computer vision \cite{chexnet17,ZSSGM18}, natural language processing \cite{GLUE,superglue} and numerous other areas \cite{ZY17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
By combining multiple information sources, it is possible to share all the information in the same model.
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
The performance of multi-task learning depends on the relationship of the information sources \cite{C97}.
When the information sources are heterogeneous, negative transfer -- where multi-task learning performs worse than single-task learning -- has often been observed \cite{AP16,BS17}.
While numerous studies have sought to alleviate negative transfer when it occurs \cite{ZY17}, a rigorous understanding to the contributing causes of negative transfer has remained elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.
In this work, we develop technical tools to compare multi-task learning to single-task learning for the setting of high-dimensional linear regression. % for learning from multiple linear regression tasks.
We use the tools to explain negative transfer and provide implications for positive transfer.
Based on the theory, we show practical implications for detecting and mitigating negative transfer.
%We consider a setting where the target task has limited labeled data and show
%On the other hand, unless the structures across task data are well-understood, applying multi-task learning on several different datasets often result in suboptimal models (or negative transfer in more technical terms).

@@

@@

We combine all the labeled data using a hard parameter sharing architecture that contains a shared body $B\in\real^{p\times r}$ for all tasks and a separate prediction head $\set{W_i \in \real^{r}}_{i=1}^t$ for each task \cite{R17,MTDNN19,WZR20}.
This corresponds to minimizing the following objective.
\begin{align}
	\label{eq_mtl}
	f(B; W_1, \dots, W_t) = \sum_{i=1}^t \norm{X_i B W_i - Y_i}^2.
\end{align}
For a target task $t$,
let $\hat{\beta}_t^{\MTL} = B W_t$ denote the optimal multi-task estimator by solving equation \eqref{eq_mtl}.
Let $\hat{\beta}_t^{\STL}$ denote the standard linear regression estimator using task $t$ alone.
We say there is negative transfer if the test error of $\hat{\beta}_t^{\MTL}$  is larger than that of $\hat{\beta}_t^{\STL}$, or positive transfer otherwise.

@@

The tradeoff that we developed in Section \ref{sec_main} depends on the qualities of each task dataset such as its data size and covariance matrix.
In Section \ref{sec_insight}, we use the technical results to provide guidance on when positive transfer is likely to occur and identify causes of negative transfer.
\squishlist
		\item First, we provide a more precise interpretation to a folklore understanding in multi-task learning, which posits that negative transfer is caused by having dissimilar tasks.
		For an example of two tasks, we show a sharp transition from positive to negative transfer determined by the ratio of $\norm{\beta_1 - \beta_2}^2$ and a certain function of data sizes (cf. Proposition \ref{prop_dist_transition}).
		We further show that positive transfer is more likely to occur when transferring from a less noisy source task to a more noisy target task.
		In Section \ref{sec_validate}, we validate the observation on text and image classification tasks.
%	In , we provide the trade-off between $\norm{\beta_1 - \beta_2}^2$ and a certain function $\Phi(\rho_1, \rho_2)$ to determine the type of transfer.
		\item Second, we find that depending on how large $\norm{\beta_1 - \beta_2}^2$ is, adding more labeled data from the source task does not always reduce the test error of $\hat{\beta}_t^{\MTL}$ (cf. Proposition \ref{prop_data_size}).
	We connect the observation to explain a key finding of Taskonomy \cite{ZSSGM18}.
	We define the \textit{data efficiency ratio} as the smallest $\alpha\in(0, 1)$ such that if we only use an $\alpha$ fraction of labeled data, then the test error of $\hat{\beta}_t^{\MTL}$ matches that of the $\hat{\beta}_t^{\STL}$ on the entire set.
	In Proposition \ref{prop_data_efficiency}, we show that the data efficiency ratio of an illustrative example is at most $\alert{\frac {1}{2\rho_t}}$.
	In Section \ref{sec_validate}, we validate that performing multi-task learning can reduce the need for labeled data on 6 sentiment analysis tasks. %, we observe that just by using \alert{40\%} of the labeled data, the overall accuracy of multi-task learning matches that of learning every task in isolation.
		\item Finally, we find that covariate shift, i.e. having different covariance matrices, is another cause for suboptimal performance for $\hat{\beta}_t^{\MTL}$.
		We show that as $n_1 / n_2$ becomes large, the best performing source task has have the same covariance matrix as the target task (cf. Proposition \ref{prop_covariate}).
%		On the other hand, when $n_1 / n_2$ is small, there are counter examples where having the same covariance matrix is not necessarily the optimal choice.
\squishend
